Deciphering enhancer rules underlying early human neuronal and neural crest development.

### 2025-04-04 manuscript.tex
continue M&M.

\subsubsection*{Validation of DeepNeuralTube organoid based on Vista enhancers}
Data was downloaded from the VISTA Enhancer browser\cite{viselVISTAEnhancerBrowsera2007, kosickiVISTAEnhancerBrowser2025} (\href{https://gitlab.com/egsb-mfgl/vista-data/-/raw/main/experiments.tsv.gz}{https://gitlab.com/egsb-mfgl/vista-data/-/raw/main/experiments.tsv.gz}) and the DNA sequence of each element was obtain either from the human genome (hg38) or the mouse genome (mm10) dependent on the origin of each sequence. In case a sequence was shorter than 500 bp (the input size of DeepNeuralTube) the sequence was extended to 500 bp based on the genomic coordinates. In case the sequence was larger than 500 bp a sliding window, with step size of 1, was produced over the sequence. A prediction score for each sequence and using both $DeepNeuralTube_O$ and $DeepNeuralTube_E$ was calculated and per experiment and output class of the model the maximum prediction score was retained. 
Ground truth neural tube enhancers were defined as enhancers that were positive in the tissue "nt", in at least 3 embryos and at
least \SI{50}{\percent} of tested embryos. 
Ground truth facial mesenchyme enhancers were defined as enhancers that were positive in the tissue "fm", in at least 3 embryos and at
least \SI{50}{\percent} of tested embryos. 
For both neural tube enhancers and facial mesenchyme enhancers precision-recall curves and receiver-operator curves were
calculated using respectively the function metrics.precision\_recall\_curve and metrics.roc\_curve from Scikit-learn
(v1.5.2; \href{https://rrid.site/data/record/nlx_144509-1/SCR_002577/resolver}{RRID:SCR\_002577}) using the maximum
prediction score across classes related to neural tube respectively facial mesenchyme in both organoid and embryo. 
We consider classes related to neural tube in the organoid as classes of the following topics: 
neural crest Topic 5,
neural crest Topic 7,
neuron Topic 1,
neuron Topic 2,
neuron Topic 3,
neuron Topic 4,
neuron Topic 6,
neuron Topic 10,
neuron Topic 11,
neuron Topic 12,
neuron Topic 13,
neuron Topic 15,
neuron Topic 16,
neuron Topic 18,
neuron Topic 19,
neuron Topic 20,
neuron Topic 21,
neuron Topic 23,
neuron Topic 24,
neuron Topic 25,
progenitor Topic 1,
progenitor Topic 3,
progenitor Topic 8,
progenitor Topic 9,
progenitor Topic 11,
progenitor Topic 13,
progenitor Topic 14,
progenitor Topic 16,
progenitor Topic 19,
progenitor Topic 21,
progenitor Topic 23,
progenitor Topic 24,
progenitor Topic 25,
progenitor Topic 29 and
progenitor Topic 30
We consider classes related to neural tube in the embryo as classes of the following topics: 
│   │   "neuron_Topic_1",  #
     7   │   │   "neuron_Topic_3",  #
     6   │   │   "neuron_Topic_5",  #
     5   │   │   "neuron_Topic_6",  #
     4   │   │   "neuron_Topic_7",
     3   │   │   "neuron_Topic_8",  #
     2   │   │   "neuron_Topic_9",  #
     1   │   │   "neuron_Topic_10",
  1016   │   │   "neuron_Topic_11",  #
     1   │   │   "neuron_Topic_12",
     2   │   │   "neuron_Topic_13",  #
     3   │   │   "neuron_Topic_14",
     4   │   │   "neuron_Topic_15",  #
     5   │   │   "neuron_Topic_17",  #
     6   │   │   "neuron_Topic_18",  #
     7   │   │   "neuron_Topic_19",  #
     8   │   │   "neuron_Topic_22",  #
     9   │   │   "neuron_Topic_24",  #
    10   │   │   "neuron_Topic_26",  #
    11   │   │   "neuron_Topic_27",  #
    12   │   │   "neuron_Topic_29",
    13   │   │   "neuron_Topic_30",
    14   │   │   "progenitor_Topic_1",  #
    15   │   │   "progenitor_Topic_3",
    16   │   │   "progenitor_Topic_4",  #
    17   │   │   "progenitor_Topic_8",  #
    18   │   │   "progenitor_Topic_10",
    19   │   │   "progenitor_Topic_11",
    20   │   │   "progenitor_Topic_16",
    21   │   │   "progenitor_Topic_17",
    22   │   │   "progenitor_Topic_21",  #
    23   │   │   "progenitor_Topic_22",  #
    24   │   │   "progenitor_Topic_28",  #
    25   │   │   "progenitor_Topic_29",  #
    26   │   │   "progenitor_Topic_31",  #
    27   │   │   "progenitor_Topic_32",  #
    28   │   │   "progenitor_Topic_36",
    29   │   │   "progenitor_Topic_40",  #
    30   │   │   "progenitor_Topic_41",  #
    31   │   │   "progenitor_Topic_44",  #
    32   │   │   "progenitor_Topic_45",
    33   │   │   "progenitor_Topic_49",  #
    34   │   │   "progenitor_Topic_51",
    35   │   │   "progenitor_Topic_57",  #
    36   │   │   "progenitor_Topic_58",  #
    37   │   │   "progenitor_Topic_59",
    38   │   │   "neural_crest_Topic_12",  #
    39   │   │   "neural_crest_Topic_13",  #
    40   │   │   "neural_crest_Topic_15",
    41   
We consider classes related to facial mesenchyme in the organod as classes of the following topics: 
neural crest Topic 3
We consider classes related to facial mesenchyme in the organod as classes of the following topics: 
neural crest Topic 1, 
neural crest Topic 19, 
neural crest Topic 21 and 
neural crest Topic 29

>>>>

\subsubsection*{Validation of DeepNeuralTube organoid based on Vista enhancers}
Data was downloaded from the VISTA Enhancer browser\cite{viselVISTAEnhancerBrowsera2007, kosickiVISTAEnhancerBrowser2025} (\href{https://gitlab.com/egsb-mfgl/vista-data/-/raw/main/experiments.tsv.gz}{https://gitlab.com/egsb-mfgl/vista-data/-/raw/main/experiments.tsv.gz}) and the DNA sequence of each element was obtain either from the human genome (hg38) or the mouse genome (mm10) dependent on the origin of each sequence. In case a sequence was shorter than 500 bp (the input size of DeepNeuralTube) the sequence was extended to 500 bp based on the genomic coordinates. In case the sequence was larger than 500 bp a sliding window, with step size of 1, was produced over the sequence. A prediction score for each sequence and using both $DeepNeuralTube_O$ and $DeepNeuralTube_E$ was calculated and per experiment and output class of the model the maximum prediction score was retained. 
Ground truth neural tube enhancers were defined as enhancers that were positive in the tissue "nt", in at least 3 embryos and at
least \SI{50}{\percent} of tested embryos. 
Ground truth facial mesenchyme enhancers were defined as enhancers that were positive in the tissue "fm", in at least 3 embryos and at
least \SI{50}{\percent} of tested embryos. 
For both neural tube enhancers and facial mesenchyme enhancers precision-recall curves and receiver-operator curves were
calculated using respectively the function metrics.precision\_recall\_curve and metrics.roc\_curve from Scikit-learn
(v1.5.2; \href{https://rrid.site/data/record/nlx_144509-1/SCR_002577/resolver}{RRID:SCR\_002577}) using the maximum
prediction score across classes related to neural tube respectively facial mesenchyme in both organoid and embryo. 
We consider classes related to neural tube in the organoid as classes of the following topics: 
neural crest Topic 5,
neural crest Topic 7,
neuron Topic 1,
neuron Topic 2,
neuron Topic 3,
neuron Topic 4,
neuron Topic 6,
neuron Topic 10,
neuron Topic 11,
neuron Topic 12,
neuron Topic 13,
neuron Topic 15,
neuron Topic 16,
neuron Topic 18,
neuron Topic 19,
neuron Topic 20,
neuron Topic 21,
neuron Topic 23,
neuron Topic 24,
neuron Topic 25,
progenitor Topic 1,
progenitor Topic 3,
progenitor Topic 8,
progenitor Topic 9,
progenitor Topic 11,
progenitor Topic 13,
progenitor Topic 14,
progenitor Topic 16,
progenitor Topic 19,
progenitor Topic 21,
progenitor Topic 23,
progenitor Topic 24,
progenitor Topic 25,
progenitor Topic 29 and
progenitor Topic 30
We consider classes related to neural tube in the embryo as classes of the following topics: 
neuron Topic 1, 
neuron Topic 3, 
neuron Topic 5, 
neuron Topic 6, 
neuron Topic 7,
neuron Topic 8, 
neuron Topic 9, 
neuron Topic 10,
neuron Topic 11, 
neuron Topic 12,
neuron Topic 13, 
neuron Topic 14,
neuron Topic 15, 
neuron Topic 17, 
neuron Topic 18, 
neuron Topic 19, 
neuron Topic 22,
neuron Topic 24, 
neuron Topic 26, 
neuron Topic 27, 
neuron Topic 29,
neuron Topic 30,
progenitor Topic 1, 
progenitor Topic 3,
progenitor Topic 4, 
progenitor Topic 8, 
progenitor Topic 10,
progenitor Topic 11,
progenitor Topic 16,
progenitor Topic 17,
progenitor Topic 21, 
progenitor Topic 22, 
progenitor Topic 28, 
progenitor Topic 29, 
progenitor Topic 31, 
progenitor Topic 32, 
progenitor Topic 36,
progenitor Topic 40, 
progenitor Topic 41, 
progenitor Topic 44, 
progenitor Topic 45,
progenitor Topic 49, 
progenitor Topic 51,
progenitor Topic 57, 
progenitor Topic 58, 
progenitor Topic 59,
neural crest Topic 12, 
neural crest Topic 13, 
neural crest Topic 15,
We consider classes related to facial mesenchyme in the organod as classes of the following topics: 
neural crest Topic 3
We consider classes related to facial mesenchyme in the organod as classes of the following topics: 
neural crest Topic 1, 
neural crest Topic 19, 
neural crest Topic 21 and 
neural crest Topic 29
